### 2024-11-01 scripts/05-exploratory_data_analysis.qmd
Added significant (featured) value selection

@@

>>>>

library(moments)

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#select significant values based on variable type of each candidate

# Define the function to select feature values based on variable type
select_feature_value <- function(data, candidate_name) {
  feature_values <- list()

  for (col_name in names(data)) {
    col_data <- data[[col_name]]

    # Numeric variables: Choose mean or median based on skewness
    if (is.numeric(col_data)) {
      # Check if there are enough valid values to compute skewness
      if (sum(!is.na(col_data)) > 1) {  # Ensure there is more than one valid value
        skew_val <- skewness(col_data, na.rm = TRUE)
        if (!is.na(skew_val) && skew_val < 1) {
          feature_values[[col_name]] <- mean(col_data, na.rm = TRUE)
        } else {
          feature_values[[col_name]] <- median(col_data, na.rm = TRUE)
        }
      } else {
        # Default to median if skewness cannot be computed
        feature_values[[col_name]] <- median(col_data, na.rm = TRUE)
      }
    }

    # Categorical variables: Choose mode (most frequent value)
    else if (is.character(col_data)) {
      feature_values[[col_name]] <- Mode(col_data)
    }

    # Boolean variables: Choose TRUE or FALSE based on which one appears more frequently
    else if (is.logical(col_data)) {
      true_count <- sum(col_data, na.rm = TRUE)
      false_count <- sum(!col_data, na.rm = TRUE)
      feature_values[[col_name]] <- if (true_count >= false_count) TRUE else FALSE
    }
  }

  # Add the candidate name to the results
  feature_values[["Candidate"]] <- candidate_name
  return(feature_values)
}

# Define a function to calculate the mode (most frequent value)
Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

# Calculate feature values for each candidate
trump_features <- select_feature_value(trump_test, "Donald Trump")
harris_features <- select_feature_value(harris_test, "Kamala Harris")

# Combine results into a single data frame for easy viewing
all_candidate_features <- bind_rows(trump_features, harris_features)

# Display the final table of feature values
print(all_candidate_features)
